\begin{itemize}[leftmargin=*]
    \item
        \textbf{AWS (Amazon Web Services)}
        Various services run by Amazon for helping to create web-based infrastructure
    \item
        \textbf{Capistrano}
        A command-line tool for automating deployment
    \item
        \textbf{Client-Side}:
        Code that runs on the user's computer. As opposed to server-side, which runs on the server.
    \item
        \textbf{Controller}:
        A piece of code that is run for a specific route, whose job it is to get/update the relevant data and return a response to the user.
    \item
        \textbf{CORS (Cross-Origin Resource Sharing)}:
        A safety feature that permits browsers to make requests to APIs on different domains.
    \item
        \textbf{Deployment}
        Getting a site onto a server
    \item
        \textbf{Foreign Key}:
        A relationship between two tables in MySQL that is enforced by the database.
    \item
        \textbf{Many-to-Many}:
        A relationship between two tables of a database where the items of table A can be linked to many items from table B, and the items from table B can be linked to many items in table A.
    \item
        \textbf{Middleware}:
        A piece of code that runs before/after the controller code to change the request/response in some way.
    \item
        \textbf{Nginx}
        A web-server
    \item
        \textbf{One-to-Many}:
        A relationship between two tables of a database where the items of table A can be linked to many items from table B, but items from table B can only be linked to one item in table A.
    \item
        \textbf{ORM (Object Relational Mapper)}:
        Allows us to access data from a database using standard objects.
    \item
        \textbf{Pivot Table}:
        A table that stores a many-to-many relationship.
    \item
        \textbf{Provisioning}:
        Setting up a server/virtual machine so that it's ready to be used.
    \item
        \textbf{Refactoring}:
        Rewriting a piece of code without changing what it does.
    \item
        \textbf{Relational Database Management Systems}:
        A database made up of tables (that represent one type of thing) and the relationships between items in those tables.
    \item
        \textbf{Resource}:
        Allows us to control the JSON output of our API.
    \item
        \textbf{Server-Side}:
        Code that runs on a server. As opposed to client-side, which runs on the user's computer.
    \item
        \textbf{SSH (Secure SHell)}:
        A command-line tool that allows you to login to and run commands on a remote machine
    \item
        \textbf{Stack}:
        A set of software that is commonly used together (e.g. the LEMP server stack)
    \item
        \textbf{Web Server}:
        A piece of software that ``serves'' files in response to a request over the internet.
\end{itemize}
